---
title: "CISC351, Research Question 1"
output: pdf_document
date: "2023-04-18"
---
### Data Preprocessing

In this research question, we look at the effect a Superhost Status has on a host. First, we will use the minimum stay, price, and number of reviews to estimate the income of a host, and their occupancy.

```{r setup, include=FALSE}
library(readr)
library(dplyr)
det_data = read_csv("det_listings.csv")  # detailed listings information aka listings.csv.gz
sum_data = read_csv("sum_listings.csv")  # summary of listing information aka listings.csv
```

```{r}
# creating a dataset to estimate the occupancy and income
superhost = det_data[, c("id", "host_id", "host_is_superhost")]

info = sum_data[, c("id", "host_id", "price", "minimum_nights", "number_of_reviews", "availability_365", "calculated_host_listings_count")]

RQ1_df = merge(info, superhost, by = "id")

RQ1_df$occupancy = RQ1_df$minimum_nights * RQ1_df$number_of_reviews

RQ1_df$income = RQ1_df$minimum_nights * RQ1_df$number_of_reviews * RQ1_df$price

# removing listings where the occupancy and income = 0
RQ1_df1 = RQ1_df[!(RQ1_df$occupancy == 0 & RQ1_df$income == 0),]
```

```{r}
superhosts = RQ1_df1[!(RQ1_df1$host_is_superhost == "FALSE"),]
hosts = RQ1_df1[!(RQ1_df1$host_is_superhost == "TRUE"),]
```

### Statistical Testing
Second, a Welch Two Sample t-test will be used to compare the means of income and occupancy

```{r}
ttest_income = t.test(superhosts$income, hosts$income, alternative = "two.sided", var.equal = FALSE)
ttest_occupancy = t.test(superhosts$occupancy, hosts$occupancy, alternative = "two.sided", var.equal = FALSE)

```

```{r}
# visualizing the occupancy
RQ1_df1$log_occupancy = log(RQ1_df1$occupancy)
boxplot(log_occupancy ~ host_is_superhost, data = RQ1_df1, 
        xlab = "Host is a Superhost", ylab = "Occupancy", 
        main = "Comparison of Log Transformed Occupancies between Regular Hosts and Superhosts")
```

```{r}
RQ1_df1$log_income = log(RQ1_df1$income)
boxplot(log_income ~ host_is_superhost, data = RQ1_df1, 
        xlab = "Host is a Superhost", ylab = "Income", 
        main = "Comparison of Log Transformed Incomes between Regular Hosts and Superhosts")
```

### Prediction Modeling
Utilizing the occupancy rate as the outcome variable, the following prediction models were created to determine the degree to which the superhost status affects the occupancy rate.
#### Looking at the the Occupancy
```{r}
superhost1 = det_data[, c(1, 13, 16, 17, 18, 19, 23, 24, 25, 26, 27, 31, 32, 33, 34, 35, 36, 37, 38, 39, 43, 51, 52, 53, 54, 63, 64, 65, 66, 67, 68, 70, 75)]
info1 = sum_data[, c("id", "price", "minimum_nights", "availability_365", "number_of_reviews", "calculated_host_listings_count")]
RQ1_df2 = merge(info1, superhost1, by = "id")

RQ1_df2 = RQ1_df2[,colSums(is.na(RQ1_df2))<nrow(RQ1_df2)]  # removing columns that are empty

RQ1_df2$occupancy = RQ1_df2$minimum_nights * RQ1_df2$number_of_reviews

RQ1_df2 = RQ1_df2[!(RQ1_df2$occupancy == 0),]

RQ1_df2 = RQ1_df2[,-c(1, 3, 5)]  # removing columns utilized to predict occupancy to remove multicollinearity and the id

```

```{r}
myfit1 = lm(occupancy ~ ., data = RQ1_df2)
summary(myfit1)

# finding the significant features 
all_feat = summary(myfit1)$coefficients[-1,4] < 0.05
important_feat = names(all_feat)[all_feat == TRUE]
for (feat in important_feat) {
  print(summary(myfit1)$coefficients[-1,4][feat])
}
```

#### Looking at the Income
```{r}
superhost1 = det_data[, c(1, 13, 16, 17, 18, 19, 23, 24, 25, 26, 27, 31, 32, 33, 34, 35, 36, 37, 38, 39, 43, 51, 52, 53, 54, 63, 64, 65, 66, 67, 68, 70, 75)]
info1 = sum_data[, c("id", "price", "minimum_nights", "availability_365", "number_of_reviews", "calculated_host_listings_count")]
RQ1_df3 = merge(info1, superhost1, by = "id")

RQ1_df3 = RQ1_df3[,colSums(is.na(RQ1_df3))<nrow(RQ1_df3)]  # removing columns that are empty

RQ1_df3$income = RQ1_df3$minimum_nights * RQ1_df3$number_of_reviews * RQ1_df3$price

RQ1_df3 = RQ1_df3[!(RQ1_df3$income == 0),]

RQ1_df3 = RQ1_df3[,-1]
```

```{r}
myfit2 = lm(income ~ ., data = RQ1_df3)
summary(myfit2)
all_feat1 = summary(myfit2)$coefficients[-1,4] < 0.05
important_feat1 = names(all_feat1)[all_feat1 == TRUE]
for (feat in important_feat1) {
  print(summary(myfit2)$coefficients[-1,4][feat])
}
```

