---
title: "CISC351, Research Question 1"
output: pdf_document
date: "2023-03-09"
---

In this research question, we look at the effect a Superhost Status has on a host. First, we will use the minimum stay, price, and number of reviews to estimate the income of a host, and their occupancy.

```{r setup, include=FALSE}
library(readr)
library(dplyr)
det_data = read_csv("det_listings.csv")  # detailed listings information aka listings.csv.gz
sum_data = read_csv("sum_listings.csv")  # summary of listing information aka listings.csv
```

```{r}
# creating a dataset to estimate the occupancy and income
superhost = det_data[, c("id", "host_id", "host_is_superhost")]

info = sum_listings[, c("id", "host_id", "price", "minimum_nights", "number_of_reviews", "availability_365", "calculated_host_listings_count")]

RQ1_df = merge(info, superhost, by = "id")

RQ1_df$occupancy = RQ1_df$minimum_nights * RQ1_df$number_of_reviews

RQ1_df$income = RQ1_df$minimum_nights * RQ1_df$number_of_reviews * RQ1_df$price

# removing listings where the occupancy and income = 0
RQ1_df1 = RQ1_df[!(RQ1_df$occupancy == 0 & RQ1_df$income == 0),]
```

To determine if there is a difference in the income or occupancy of a superhost compared to a regular host, a two-sample t-test will be used to determine the mean difference in the income and occupancy. First, two new subsets will be created to distinguish the superhosts from the regular hosts.

```{r}
superhosts = RQ1_df1[!(RQ1_df1$host_is_superhost == "FALSE"),]
hosts = RQ1_df1[!(RQ1_df1$host_is_superhost == "TRUE"),]
```

Second, a Welch Two Sample t-test will be used to compare the means of income and occupancy

```{r}
occupancy_ttest = t.test(superhosts$occupancy, hosts$occupancy, alternative = "two.sided", var.equal = FALSE)
income_ttest = t.test(superhosts$income, hosts$income, alternative = "two.sided", var.equal = FALSE)
```

```{r}
occupancy_ttest
```

From comparing the mean occupancy of a superhost to a host, we can observe that there is a statistically significant difference among the guests both hosts will have. On average, superhosts will house 940 guests, while regular hosts will house 681 guests.

```{r}
income_ttest
```

From comparing the mean income of a superhost to a host, we can observe that there is not a statistically significant difference among the two incomes, though there is a slight difference with superhosts receiving an income of roughly \$20,000 more.

Thus, we can conclude that while there is not a statistically significant difference in the income of a regular host to a superhost, superhosts benefit from the superhost status as they are more likely to have more guests.
